---
title: "National Parks and `ggplot2`"
author: "Our team!"
date: "8/22/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

***

## The plan

We'll be [mob programming](https://github.com/R-Meetup-SB/hackathon-201806/blob/master/mob-programming.Rmd) in small teams to learn ggplot2, rotating who codes on a single computer so that everyone can contribute verbally and get a chance to type. Each team will have mixed skill-levels, and the idea is to learn together. Everyone can learn something new by coding with other people! Please be respectful and supportive first and foremost.  

We'll be exploring data on visits to US National Parks (sourced from the [National Park Service](https://irma.nps.gov/Stats/SSRSReports)).  

Often, you will need to do a bit of "data wrangling" to isolate, summarize, or combine data before you can make the plot that you'd like. Since data wrangling is not the focus of this Meetup, we have provided wrangling code to get you started for each plot that your group will make.

References to help you: 

- [RLadiesSB intro presentation](https://rladies.github.io/meetup-presentations_santabarbara/2018_08_22_mobprogramming/mobprogramming_and_ggplot2.html#1)
- [data visualization cheatsheet](https://www.rstudio.com/wp-content/uploads/2015/03/ggplot2-cheatsheet.pdf) 
- [ggplot reference](https://ggplot2.tidyverse.org/reference/)
- [data visualisation](http://r4ds.had.co.nz/data-visualisation.html) and [graphics for communication](http://r4ds.had.co.nz/graphics-for-communication.html) chapters in [R for data science](http://r4ds.had.co.nz/)
- help pages for the function (type `?function_name` in the console)

*Created by Julie Lowndes, Ellie Campbell, Iwen Su, and Camila Vargas*

***

## Load libraries and data 

The `tidyverse` package has `ggplot2`, as well as `dplyr`, and we will use both today. Run this code chunk to load the `tidyverse` package (install it if necessary), and read in the data as the `np_visit` object.


```{r load tidyverse and data, message=FALSE, warning=FALSE}
library(tidyverse) # install.packages('tidyverse')

np_visit <- read_csv("data/np_visit.csv")
```

***

## Demo code

This is the demo code that we went over in the presentation. 

```{r demo plot, warning=FALSE, message=FALSE}
dino_nm <- np_visit %>% 
  filter(park_name == "Dinosaur National Monument")
  
ggplot(data = dino_nm, aes(x = year, y = visitors)) +
  geom_point()
```

***

## Explore the data

When you first get your hands on a dataset, it's always good to quickly explore the contents. This will give you a better idea of what you can plot and what visuals might be interesting. Preliminary exploration may involve looking at the first few rows of the data table, identifying the type of data collected, or checking out max and min values. Individual columns can be pulled out using the notation: `dataframe$column_name`.  

Here are some common functions used for initial exploration: 

- View()
- head()
- tail()
- summary()

Raw data is rarely prepared in a format that is ready-to-go for plotting. We often need to 'wrangle' (or rearrange and tidy) the rows and columns in a way that makes sense for `ggplot` or any other analyses for that matter. 

Helpful functions from [`dplyr`](https://dplyr.tidyverse.org/) (also part of the `tidyverse`) that assist with wrangling:

- filter()
- select()
- %>%

***

## Plotting tasks

In the next section you are going to find different questions to be answered by visualizing the data in a plot. The code for wrangling the data is provided at the beginning of each code chunk. Your task is to write code and create plots and have fun with it. Here we provide questions and starter code, but feel free to follow the interests of your group and explore different capabilities of `ggplot2`! 

<br>

### How have visits to California National Parks changed over time?

Plot California National Parks over time, and distinguish each park by color. Try this as both a `geom_point` and a `geom_line`.

Hint: you can control aesthetics within the geoms: `geom_point(aes(color = park_name))`.

```{r ca parks by color}
## wrangling code:
np_ca <- np_visit %>% 
  filter(state == "CA" & type == "National Park") %>% 
  arrange(park_name, year)
```


### What is the median visitations to National Parks in California? What are the max and min values?

Visualize visitation to National Parks in California, clearly showing the median and the max and min values. Keep in mind that the type of visualization you create must be compatible with the data you use.

Hint: `geom_boxplot`  

```{r average visitation to ca parks}
## wrangling code:
np_ca <- np_visit %>%
  filter(state == "CA" & type == "National Park") %>% 
  arrange(park_name, year)

## plotting code:

```
  
**Bonus**: try `geom_jitter`! Does it help you see anything differently than boxplot did?


### How many people visit a National Park in 2016? How many visited a National Monument?

Plot visitor counts from 2016 for National Parks and Monuments. 

Hint: `geom_bar`

```{r counts from 2016}
## wrangling code:
visit_16 <- np_visit %>% 
  filter(year == 2016) %>% 
  filter(type == "National Park" | type == "National Monument")

## plotting code:

```

**Bonus**: Now color them by region: (hint: `fill = region`)  

**Bonus 2**: Control whether they are stacked (default) or not: (hint: `position = "dodge"`)


### How many parks are in each region?

The "region" variable divides the US in 8 regions. Use a bargraph to show how many parks are in each region. 

Hint: `geom_bar`

```{r parks by region}
## wrangling code:
per_region <- np_visit %>% 
  group_by(region) %>% 
  mutate(n_parks = n()) %>% 
  select(region, n_parks) %>% 
  distinct() %>% 
  arrange(desc(n_parks)) %>% 
  ungroup()

## plotting code:

```

### Compare change over time for all National Parks in California

If we want to compare visitation plots of all National Parks in California, hopefully there is a better way than creating separate figures for each one. One of the best approaches to coding is expecting that someone before you has wanted to do the same thing, so you don't need to reinvent the wheel. In `ggplot2`, can use `facet_wrap` to create separate plots by a variable that we select (in this example, park_name).

Hint: `facet_wrap(~park_name)`

```{r multiple graphs with facet_wrap}
## wrangling code:
np_ca <- np_visit %>%
  filter(state == "CA" & type == "National Park") %>% 
  arrange(park_name, year)

## plotting code:

```

**Bonus**: What does `facet_grid` do?

***

## Styling: labels, titles, and themes

`ggplot() + geom()` creates the base graphic. To customize graphs, you can add (using the plus sign) more elements such as:

- xlab()
- ylab()
- ggtitle()

One way to make major changes to the overall aesthetic of your graph is with themes (that you also add using `+`). Themes to try:

- theme_bw()
- theme_minimal()
- theme_classic()

```{r customize demo plot}
ggplot(data = dino_nm, aes(x = year, y = visitors)) +
  geom_point() +
  xlab("Year") +
  ylab("Annual Visitors") +
  ggtitle("California National Parks Visitation") +
  theme_bw()
```

Finally, if you still have time, explore the many arguments of [`theme`](https://ggplot2.tidyverse.org/reference/theme.html) which can be used to more tightly control theme elements. And try customizing any of the plots created above (eg: change the color scheme, add titles and labels, change the position of the key, etc.) 

***
