---
title: "National Parks Analysis"
author: "Us!"
date: "8/22/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## The plan

We'll be mob programming in small teams to learn ggplot2. We'll be exploring data on visits to US National Parks (sourced from the National Park Service at https://irma.nps.gov/Stats/SSRSReports).  



## Load libraries and data 

The `tidyverse` package has `ggplot2`, as well as `dplyr`, which we will use today.

```{r load, message=FALSE, warning=FALSE}
library(tidyverse) # install.packages('tidyverse')

np_visit <- read_csv("data/np_visit.csv")
```


## Exploring and wrangling data, pre-plotting

To get a better sense of the dataset and what we may want to plot, we can do some preliminary wrangling/exploration.


```{r}

## General exploration of the dataset
head(np_visit)
summary(np_visit)

## How many years of data does this dataset have? hint: summary()

## Which parks, and how many parks, are in the dataset?
parks <- np_visit %>% 
  select(park_name) %>% 
  distinct()

## When was each park founded? Which are the oldest vs newest?
earliest_yr <- np_visit %>% 
  select(park_name, year) %>% 
  group_by(park_name) %>% 
  slice(which.min(year)) %>%
  arrange(year) %>% 
  ungroup()

head(earliest_yr)
tail(earliest_yr)

## How many parks in each state? Each region?
per_state <- np_visit %>% 
  group_by(state) %>% 
  mutate(n_parks = n()) %>% 
  select(state, n_parks) %>% 
  distinct() %>% 
  arrange(desc(n_parks)) %>% 
  ungroup()

head(per_state)
tail(per_state)

per_rgn <- np_visit %>% 
  group_by(region) %>% 
  mutate(n_parks = n()) %>% 
  select(region, n_parks) %>% 
  distinct() %>% 
  arrange(desc(n_parks)) %>% 
  ungroup()

## Which is the most visited national park? Which year? How many visitors?
most_visit <- np_visit %>% 
  arrange(desc(visitors))

## Which was the most visited national park in the US in 2016?
most_visit_2016 <- np_visit %>% 
  filter(year == 2016) %>% 
  arrange(desc(visitors))

## And the least visited park in 2016
tail(most_visit_2016)


```


## Demo plots


```{r}
ggplot(data = earliest_yr, aes(year)) + geom_histogram(binwidth = 1) # single year
ggplot(data = earliest_yr, aes(year)) + geom_histogram(binwidth = 10) # decade

rgn_visits <- np_visit %>% 
  group_by(region, year) %>% 
  mutate(rgn_visit = sum(as.numeric(visitors))/10^4) %>% 
  ungroup() %>% 
  filter(rgn_visit > 0)

ggplot(data = rgn_visits, aes(x = year, y = rgn_visit, colour = region)) + 
  geom_line()
```

## Your turn!

Here are some questions to get you started: 

1. Can you create a plot that shows the trend in visitation to your favorite national park over time
2. How many national parks were established in each year in the US between XXXX and XXXX
3. 
